\todo{Le titre}

### 2013-03-11 parts/def-ph.tex
Plus de formalisme

\towrite{Reformuler (actuellement trop proche de CMSB)}
A PH (\pref{def:PH}) gathers a finite number of concurrent \emph{processes} grouped into a finite set of \emph{sorts}.
A process belongs to a unique sort and is noted $a_i$ where $a$ is the sort and $i$ the identifier of the process within the sort $a$.
At any time, one and only one process of each sort is present; a state of the PH thus corresponds to the set of such processes.

@@

\begin{comment}
\begin{definition}[Process Hitting]\label{def:PH}
A \emph{Process Hitting} is a triple $(\PHs,\PHl,\PHa)$:
\begin{itemize}
\item $\PHs \DEF \{a,b,\dots\}$ is the finite set of \emph{sorts};
\item $\PHl \DEF \prod_{a\in\PHs} \PHl_a$ is the set of states with $\PHl_a = \{a_0,\dots,a_{l_a}\}$
the finite set of \emph{processes} of sort $a\in\Sigma$ and $l_a$ a positive integer with
	$a\neq b\Rightarrow \forall(a_i,b_j)\in\PHl_a\times\PHl_b,a_i\neq b_j$;
\item $\PHa \DEF \{ \PHfrappe{a_i}{b_j}{b_k}, \dots \mid
					(a,b)\in\PHs^2 \wedge (a_i,b_j,b_k)\in \PHl_a\times\PHl_b\times\PHl_b$ \\
	\hspace*{2cm} $\wedge b_j\neq b_k \wedge a=b\Rightarrow a_i=b_j\}$
			is the finite set of \emph{actions}.
\end{itemize}
$\PHproc$ denotes the set of all processes ($\PHproc \DEF \{ a_i\mid a\in\PHs \wedge a_i\in\PHl_a\}$).
\end{definition}
\end{comment}

@@

@@

\paragraph{Modeling cooperation.}
\towrite{Sortes coopératives avec mise à jour prioritaire (pas de décalage temporel)}
As described in \cite{PMR10-TCSB}, the cooperation between processes to make another bounce can be
expressed in PH by building a \emph{cooperative sort}.
\pref{fig:PH-cooperativity} shows an example of cooperation between processes $b_1$ and $c_1$ to
make $a_1$ bounce to $a_2$:
a cooperative sort $bc$ is defined with 4 processes (one for each sub-state of the presence of
processes $b_1$ and $c_1$).
For the sake of clarity, the $bc$ processes are indexed using the sub-state they represent.
Hence, $bc_{01}$ represents the sub-state $\PHstate{b_0,c_1}$, and so on.
Each process of sort $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the status of the sorts $b$
and $c$ (e.g., $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).
Then, it is the process $bc_{11}$ which hits $a_1$ to make it bounce to $a_2$ instead of the
independent hits from $b_1$ and $c_1$.

@@

While the construction of cooperation in PH allows to encode any boolean functions
between cooperating processes \cite{PMR10-TCSB}, it is worth noticing they introduce a temporal
shift in their application.
This allows the existence of interleaving of actions leading to a cooperative sort representing a
past sub-state of the presence of the cooperative processes.
The resulting behavior is then an over-approximation
of the realization of an instantaneous cooperation.

@@

\todo{Modèles non-Zeno (OU : sans domaines-pièges)}

>>>>

\subsection{Definition of the Process Hitting with $k$ priorities}
\towrite{Reformuler (actuellement trop proche de CMSB)}
A PH (\pref{def:PH}) gathers a finite number of concurrent \emph{processes} grouped into a finite set of \emph{sorts}.
A process belongs to a unique sort and is noted $a_i$ where $a$ is the sort and $i$ the identifier of the process within the sort $a$.
At any time, one and only one process of each sort is present; a state of the PH thus corresponds to the set of such processes.

@@

% Process Hitting sans priorités
\begin{comment}
\begin{definition}[Process Hitting]\label{def:PH}
A \emph{Process Hitting} is a triple $(\PHs,\PHl,\PHa)$:
\begin{itemize}
\item $\PHs \DEF \{a,b,\dots\}$ is the finite set of \emph{sorts};
\item $\PHl \DEF \prod_{a\in\PHs} \PHl_a$ is the set of states with $\PHl_a = \{a_0,\dots,a_{l_a}\}$
the finite set of \emph{processes} of sort $a\in\Sigma$ and $l_a$ a positive integer with
	$a\neq b\Rightarrow \forall(a_i,b_j)\in\PHl_a\times\PHl_b,a_i\neq b_j$;
\item $\PHa \DEF \{ \PHfrappe{a_i}{b_j}{b_k}, \dots \mid
					(a,b)\in\PHs^2 \wedge (a_i,b_j,b_k)\in \PHl_a\times\PHl_b\times\PHl_b$ \\
	\hspace*{2cm} $\wedge b_j\neq b_k \wedge a=b\Rightarrow a_i=b_j\}$
			is the finite set of \emph{actions}.
\end{itemize}
$\PHproc$ denotes the set of all processes ($\PHproc \DEF \{ a_i\mid a\in\PHs \wedge a_i\in\PHl_a\}$).
\end{definition}
\end{comment}

@@

\begin{definition}[Process Hitting $n$-reduction] \towrite{Revoir terminologie}
  If $\PH = (\PHs; \PHl; \PHa^{\langle k \rangle})$ is a Process Hitting with $k$ priorities and $n \in \segm{1}{k}$, we denote $\restriction{\PH}{n}$ \towrite{$\PH_{|n}$ $\PH'$} the $n$-reduction of $\PH$ the following Process Hitting with $n$ priorities, where:
  $$\restriction{\PH}{n} = (\PHs; \PHl; \PHa^{\langle n \rangle})$$
  and:
  $$\PHa^{\langle n \rangle} = (\PHa^{(1)}; \dots; \PHa^{(n)})$$
\end{definition}

@@

\subsection{Modeling cooperation.}
\towrite{Sortes coopératives avec mise à jour prioritaire (pas de décalage temporel)}
As described in \cite{PMR10-TCSB}, the cooperation between processes to make another bounce can be
expressed in PH by building a \emph{cooperative sort}.
\pref{fig:PH-cooperativity} shows an example of cooperation between processes $b_1$ and $c_1$ to
make $a_1$ bounce to $a_2$:
a cooperative sort $bc$ is defined with 4 processes (one for each sub-state of the presence of
processes $b_1$ and $c_1$).
For the sake of clarity, the $bc$ processes are indexed using the sub-state they represent.
Hence, $bc_{01}$ represents the sub-state $\PHstate{b_0,c_1}$, and so on.
Each process of sort $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the status of the sorts $b$
and $c$ (e.g., $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).
Then, it is the process $bc_{11}$ which hits $a_1$ to make it bounce to $a_2$ instead of the
independent hits from $b_1$ and $c_1$.

@@

While the construction of cooperation in PH allows to encode any boolean functions
between cooperating processes \cite{PMR10-TCSB}, it is worth noticing they introduce a temporal
shift in their application. \towrite{Adapter avec priorités}
This allows the existence of interleaving of actions leading to a cooperative sort representing a
past sub-state of the presence of the cooperative processes.
The resulting behavior is then an over-approximation
of the realization of an instantaneous cooperation.

@@

\subsection{Restrictions} \towrite{Reformuler}

In this section, we describe several restrictions.
Although the definition of a Process Hitting with $k$ priorities allows a wide range of models, the method given in the next section only applies to Process Hitting models that regard these restrictions.

Let $\PH$ be a PH model on which we want to apply the static analysis. This PH has to regard the two restrictions below.

\paragraph{Cycle-freeness of $\restriction{\PH}{k-1}$}
\towrite{Définition formelle ? + justification}
The model $\restriction{\PH}{k-1}$ contains no cycles. Cycles are still allowed in $\PH$ however.

\paragraph{Local priorities}
\towrite{Justification}
All actions hitting the same sort belong to the same class of priority:
$$\forall h_1, h_2 \in \PHh, \PHtarget(h_1) = \PHtarget(h_2) \Rightarrow \prio(h_1) = \prio(h_2)$$
Furthermore, for all sort $a$, we denote $\prio(a)$ the priority of the actions hitting $a$, if any:
$$\forall a \in \PHs, \exists h \in \PHh, \PHtarget(h) = a \Rightarrow \prio(a) = \prio(h)$$
\towrite{Ou : $\nexists h \Rightarrow \prio(a) = 0$}

### 2013-03-11 parts/sa.tex
Plus de formalisme

\newcommand{\abstr}[1]{#1^\wedge}%\text{\textasciicircum}}
\def\BS{\mathbf{BS}}
%\def\priomax{\mathsf{prio}_{max}}
\def\procs{\mathsf{procs}}
\def\allprocs{\mathsf{allProcs}}
\def\pfp{\mathsf{pfp}}

@@

%Plus haute priorité :
%$$\forall a \in \PHs, \priomax(a) = \max_{h \in \PHh, \PHtarget(h) = a}(\prio(h))$$

Processus possibles :
\begin{align*}
\procs((\cwSol,\cwReq,\cwCont)) = \{ p \in \PHproc \mid & \exists (P,ps) \in \cwSol, p \in ps
\\ & \vee p = \PHtarget(P)
\\ & \vee (P \neq \omega \Rightarrow p = \PHbounce(P)) \}
%\\ & \vee \exists h \in \BS(P), (p = \PHhitter(h) \vee p = \PHbounce(h)) \}
\end{align*}

Processus rencontrés :
\begin{align*}
\allprocs((\cwSol,\cwReq,\cwCont)) = \{ p \in \PHproc \mid & \exists (P,ps) \in \cwSol, p \in ps
\\ & \vee p = \PHtarget(P)
%\\ & \vee (P \neq \omega \Rightarrow p = \PHbounce(P)) \}
\\ & \vee \exists h \in \BS(P), (p = \PHhitter(h) \vee p = \PHbounce(h)) \}
\end{align*}

@@

Points fixes possibles : pour toute sorte $a$ et tout contexte partiel $\ctx$ sur $V^n(a)$:
\begin{align*}
  \pfp_\ctx: \PHs &\rightarrow \wp(\PHproc) \\
  a &\mapsto \{ s \in \times_{b \in V^n(a)} \PHl_b \mid \text{$s$ est accessible depuis $\ctx$ et aucune action n'y est jouable} \}
\end{align*}

Points fixes possibles sur un ensemble de sortes :
\begin{align*}
  \pfp: \mathbb{A} &\rightarrow \wp(\PHproc) \\
  \myB &\mapsto \bigcup_{a \in A} \pfp_{\allprocs(\myB)}(a)
\end{align*}

@@

\subsection{Abstract structure}
Séquences de bonds abstraites :
$$\BS^\wedge(P) = \{ \zeta^\wedge \mid \zeta \in \BS(P), \nexists \zeta' \in \BS(P), \zeta'^\wedge \subsetneq \zeta^\wedge \}$$
where:
$$\zeta^\wedge = \{ \PHhitter(\zeta_n) \mid n \in \indexes{\zeta} \wedge \PHsort(\PHhitter(\zeta_n)) \neq \PHsort(P) \}$$

The abstract structure $\cwB=(\Breq,\Bsol,\Bcont)$ is defined as
$
 \cwB = \sfp{\myB}{\myB}{\aB^\w_{\ctx \Cap (\procs(\myB) \cup \pfp(\myB))}}
$,
with $\myB=(\myreq,\mysol,\mycont)$:
\begin{align*}
\myreq = &\{ (a_i,\PHobjp{a}{j}{i}) \in \PHproc \times \Obj \mid a_j \in \PHget{\ctx}{a}
  \wedge (\exists (P,ps) \in \mysol, a_i \in ps \\
& \qquad\qquad
          \vee \exists n \in \indexes{\w}, \PHbounce(\w)=a_i\}
\\
\mysol \subseteq & \{ (P,ps) \in \Obj \times \powerset(\PHproc) \mid
        \exists (a_i,P)\in \myreq \wedge ps \in \aBS(P)  \}
\\
\mycont = & \{ (P,\PHobj{q}{\PHbounce(P)}) \in \Obj \times \Obj \mid 
        \exists (P,ps)\in \mysol \\
&\qquad\qquad
        \wedge q \in \gCont_\ctx(\PHs(P),P)\}
\end{align*}

>>>>

\newcommand{\abstr}[1]{#1^\wedge}%\text{\textasciicircum}}
\def\BS{\mathbf{BS}}
%\def\priomax{\mathsf{prio}_{max}}
\def\procs{\mathsf{procs}}
\def\allprocs{\mathsf{allProcs}}
\def\pfp{\mathsf{pfp}}
\def\pfpprocs{\mathsf{pfpProcs}}

@@

%%% N'est plus utile (?)
%Plus haute priorité :
%$$\forall a \in \PHs, \priomax(a) = \max_{h \in \PHh, \PHtarget(h) = a}(\prio(h))$$

Processus possibles :
\begin{align*}
\procs((\cwSol,\cwReq,\cwCont)) = \{ p \in \PHproc &\mid \exists (P,ps) \in \cwSol, p \in ps
\\ & \vee p = \PHtarget(P)
\\ & \vee (P \neq \omega \Rightarrow p = \PHbounce(P)) \}
%\\ & \vee \exists h \in \BS(P), (p = \PHhitter(h) \vee p = \PHbounce(h)) \}
\end{align*}

Processus rencontrés :
\begin{align*}
\allprocs((\cwSol,\cwReq,\cwCont)) = \{ p \in \PHproc &\mid \exists (P,ps) \in \cwSol, p \in ps
\\ & \vee p = \PHtarget(P)
%\\ & \vee (P \neq \omega \Rightarrow p = \PHbounce(P)) \}
\\ & \vee \exists h \in \BS(P), (p = \PHhitter(h) \vee p = \PHbounce(h)) \}
\end{align*}

@@

\towrite{Formaliser}

%%% Pas utile
%Points fixes possibles : pour toute sorte $a$ et tout contexte partiel $\ctx$ sur $V^n(a)$:
%\begin{align*}
%  \pfp_\ctx: \PHs &\rightarrow \wp(\PHproc) \\
%  a &\mapsto \{ s \in \underset{b \in V^n(a)}{\times} \PHl_b \mid \text{$s$ est accessible depuis $\ctx$ et aucune action n'y est jouable} \}
%\end{align*}

Points fixes possibles : pour toute sorte $a$ et tout contexte $\ctx$:
\begin{align*}
  \pfp_\ctx: \PHs &\rightarrow \wp(\PHproc) \\
  a &\mapsto \{ s \in \underset{b \in V^n(a)}{\times} \PHl_b \mid \text{$s$ est (localement) accessible dans $\restriction{\PH}{\prio(a)}$ depuis $\ctx$}\\
  &\qquad\qquad\qquad\qquad \text{et aucune action n'y est jouable} \}
\end{align*}

Processus rencontrés comme résultats d'un point fixe :
\begin{align*}
\pfpprocs_\ctx(a) = \{ \PHget{s}{a} \mid s \in \pfp_\ctx(a) \}
\end{align*}

%%% N'est plus adapté
%Points fixes possibles sur un ensemble de sortes :
%\begin{align*}
%  \pfp: \mathbb{A} &\rightarrow \wp(\PHproc) \\
%  \myB &\mapsto \bigcup_{a \in A} \pfp_{\allprocs(\myB)}(a)
%\end{align*}

@@

\subsection{Abstract structure}
%Séquences de bonds abstraites :
%$$\BS^\wedge(P) = \{ \zeta^\wedge \mid \zeta \in \BS(P), \nexists \zeta' \in \BS(P), \zeta'^\wedge \subsetneq \zeta^\wedge \}$$
%where:
%$$\zeta^\wedge = \{ \PHhitter(\zeta_n) \mid n \in \indexes{\zeta} \wedge \PHsort(\PHhitter(\zeta_n)) \neq \PHsort(P) \}$$

The abstract structure $\cwB=(\Breq,\Bsol,\Bcont)$ is defined as
$
 \cwB = \sfp{\myB}{\myB}{\aB^\w_{\ctx \Cap \allprocs(\myB)}}
$,
with $\myB=(\myreq,\mysol,\mycont)$:
\begin{align*}
\myreq &= \{ (a_i,\PHobjp{a}{j}{i}) \in \PHproc \times \Obj \mid
  (a_j \in \PHget{\ctx}{a} \vee a_j \in \pfpprocs_\ctx(a)) \\
  & \qquad \wedge (\exists (P,ps) \in \mysol, a_i \in ps \vee \exists n \in \indexes{\w}, \PHbounce(\w_n)=a_i) \}
\\
\mysol &\subseteq \{ (P,ps) \in \Obj \times \powerset(\PHproc) \mid
        \exists (a_i,P)\in \myreq \wedge ps \in \aBS(P) \}
\\
\mycont & = \{ (P,\PHobj{q}{\PHbounce(P)}) \in \Obj \times \Obj \mid 
        \exists (P,ps)\in \mysol \\
&\qquad\qquad
        \wedge q \in \gCont_\ctx(\PHs(P),P)\}
\end{align*}
